---
title: "Subset original data and append river temperature data"

---

This code is was used to produce the subset dataset used for each model which can be found in the data-raw/ folder. 

This data is not necessary to run, but kept for reference.


```{r load_libraries, eval=FALSE}
library(data.table)
library(here)
library(tidyverse)
library(zoo)
```


# Original dataset

Original data sets were shared by JG via baloo link  - included in folder as zip file for reference. Using code below a subset of .csv per species and rear type was created and saved in the data-raw/HydroSurvDOYTEMP_data_subset.zip folder. 

```{r wrangle_orig_data,  eval=FALSE}


# DATA <- read.csv(file = here("Examples/data", "dsynth_2021-03-02.spsu.subset.csv"), stringsAsFactors = FALSE, na.strings = c("", "NA"))
# 
# DATA <- read.csv(file = here("Examples/data", "dsynth_2021-03-02.stl.subset.csv"), stringsAsFactors = FALSE, na.strings = c("", "NA"))

DATA <- read.csv(file = file.choose(), stringsAsFactors = FALSE, na.strings = c("", "NA")) #select either spsu  ch or stl subset file

df<- DATA %>%
  # filter to include wild all run types, with a transport code released from LGR
  filter(
    t_rear_type == "W", #adjust as needed
    !is.na(pass_type),
    str_detect(rel_site, "LGR")
  ) %>%
  # extract year/doy and filter to include 80 to 160 doy from 1993 to 2018
  mutate(
    year = lubridate::year(rel_time),
    doy = lubridate::yday(rel_time),
  ) %>%
  
  filter(
    between(year, 1993, 2024),
    between(doy, 80, 160)
  ) %>%
  # redesignate tranportation codes--double check on -S and -T at other locations 
  mutate(pass_type_T_R = case_when(
    str_detect(pass_type, "-S") & str_length(pass_type) == 5 ~ "remove",
    str_detect(pass_type, "-S") & str_length(pass_type) > 5 ~ "ROR",
    str_detect(pass_type, "-TB") ~ "ROR",
    str_detect(pass_type, "-TD") ~ "ROR",
    str_detect(pass_type, "ROR") ~ "ROR",
    str_detect(pass_type, "LWG-T") ~ "T",
    .default = "remove"
  )) %>%
  filter(pass_type_T_R != "remove") %>% 
  # # adding filters to match JG shared code for transport --- check this still stands
  # filter(if_else(pass_type_T_R == "ROR", 
  #                str_length(pass_type) %in% c(3, 7),
  #                str_detect(pass_type, "LWG-T")
  #                )
  #        ) %>%
  # add binary codes
  mutate(
    transport = if_else(pass_type_T_R == "T", 1, 0),
    alive = if_else(!is.na(adu_lgr_first), 1, 0),
    doyz = scale(doy, center = TRUE, scale = TRUE),
    year = as.factor(year)
  ) 
```

# River data

This code is used to import LGR river temperature data and calculate 7-day rolling averages for the years 1993-2018. The data is then merged into a larger dataset for further analysis. 


```{r setup, eval=FALSE}

#set years
years<-c(1993:2020,2022:2024) #drop 2021 LWG to be replaced with tailrace
# Concatenate the years into the URL
year_param <- paste(sprintf("year%%5B%%5D=%s", years), collapse = "&")

# Create the URL with the updated year parameter and multiple sites
url <- sprintf("https://www.cbr.washington.edu/dart/cs/php/rpt/mg.php?sc=1&mgconfig=river&outputFormat=csvSingle&%s&loc%%5B%%5D=LWG&data%%5B%%5D=Temp+%%28WQM%%29&startdate=1%%2F1&enddate=12%%2F31&avgyear=0&consolidate=1&grid=1&y1min=0&y1max=&y2min=&y2max=&size=large", year_param)

lwg_forebay_temp <- read.csv(url, header = TRUE, stringsAsFactors = FALSE) %>%
  drop_na() %>% 
  mutate(date = ymd(paste(year, mm.dd, sep = "-")),
         doy = yday(date)
  )

#pull in 2021 tailrace lwg temperature -- missing data
url_2021<-"https://www.cbr.washington.edu/dart/cs/php/rpt/mg.php?sc=1&mgconfig=river&outputFormat=csvSingle&year%5B%5D=2021&loc%5B%5D=LGNW&data%5B%5D=Temp+%28WQM%29&startdate=1%2F1&enddate=12%2F31&avgyear=0&consolidate=1&grid=1&y1min=0&y1max=&y2min=&y2max=&size=large"

# wrangle 2021 data to append to historical data
lnwg_2021_temp<-read.csv(url_2021, header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(date = ymd(paste(year, mm.dd, sep = "-")),
         doy = yday(date)
         ) %>% 
  drop_na()
#combine 2021 with rest of years
LWG.temp.data<- lwg_forebay_temp %>%
  bind_rows(lnwg_2021_temp)

LWG.temp.data <- LWG.temp.data %>%
  select(year,doy,value) %>% 
  arrange(year, doy) %>%
  mutate(mean.temp = c(rep(NA,6), zoo::rollmean(value, k = 7, align = "right"))) %>% 
  select(-value) %>% 
  mutate(year = as.factor(year))


```

```{r bind_temp_data, eval=FALSE}

#merge with subset dataset from code chunk above (wrangle_orig_data) or import already saved subset file in data-raw/ 
# df<-file.choose() #choose the subsetted dataset to merge with river temperature data 


df<- df%>% 
  left_join(y= LWG.temp.data, by = c("year", "doy")) %>% 
  mutate ( mean.tempz = scale(mean.temp, center = TRUE, scale = TRUE)
  ) 

# save subset output to data-raw/ folder and update naming
write.csv(df, row.names = F, here("data-raw/HydroSurvDOYTEMP_data_subset", "natural_steelhead_data_subset.csv"))
```

