---
title: "HydroSurvDOYTEMP_brms_models"
output: html_notebook
---


Below outlines code used to run each model save in data-raw/models/ folder. Not necessary to run application, but kept for reference

```{r packages}
library(brms)
library(here)
library(tidyverse)
library(knitr)
```




# Model 
# loop model for each species and rear type
use to loop through all data_subset files and run models for each species and rear type. Save to data-raw/models folder

```{r fct_run_models}

run_models<- function(filename){
  
#import saved subset file 
df<-read_csv(here("data-raw/HydroSurvDOYTEMP_data_subset", filename))

#for binomial left hand trial 
df.agg <- df%>% 
  #calculate # of juveniles--used as total trials in binomial model
  group_by(year, doy, doyz, transport) %>% 
  mutate(doy = as.numeric(doy),
         n = n()) %>% 
  arrange(year, doy, doyz) %>% 
 # mutate(year = as.factor(year)) %>% 
  #set reponse as sum of alive per trials 
  summarize(alive = sum(alive),
            n = mean(n),
            mean.tempz = mean(BON.tempz),
            mean.temp = mean(BON.temp)) %>% 
  ungroup() %>% 
  #include sar based on pit tag returns of observed data-- binned by year, weekly doy (doy.bin), and transport (ROR v Transport)
  arrange(year, doy, doyz, transport) %>% 
  mutate(doy.bin = cut(doy, breaks = seq(90, 160, 10), include.lowest = TRUE)) %>% 
  group_by(year, transport, doy.bin) %>% 
  mutate(doy.median = round(median(doy), 0),
         sar.pit = alive/n) %>% 
         # sar.pit.se = sd(sar.pit)/sqrt(n),
         # sar.pit.lower = mean(sar.pit) - 1.96 * sar.pit.se,
         # sar.pit.upper = mean(sar.pit) + 1.96 * sar.pit.se) %>% 
  ungroup() %>% 
  mutate( date = parse_date_time(x = paste(year, doy), orders = "yj")) %>% 
  mutate(year = as.character(year),
         transport = as.character(transport))

bprior <- c(prior(normal(0, 5), class = b), 
            prior(normal(0, 5), class = Intercept),
            prior(normal(0, 1), class = sd),
            prior(normal(0, 1), class = sd, group = year))

mod_doy <- brm(alive | trials(n) ~ doyz + I(doyz^2) + transport + doyz:transport + I(doyz^2):transport + (1 + doyz + I(doyz^2) + transport | year), 
                   data = df.agg, 
                   family = binomial(link = "logit"), 
                   chains = 3,
                   cores = 3,
                   warmup = 1000, 
                   iter = 2000, 
                   thin = 1,
                   prior = bprior, 
                   seed = 123, 
                   init = "0"
                   )

mod_temp <- brm(alive | trials(n) ~ mean.tempz + I(mean.tempz^2) + transport + mean.tempz:transport + I(mean.tempz^2):transport + (1 + mean.tempz + I(mean.tempz^2) + transport | year), 
                   data = df.agg, 
                   family = binomial(link = "logit"), 
                   chains = 3,
                   cores = 3,
                   warmup = 1000, 
                   iter = 2000, 
                   thin = 1,
                   prior = bprior, 
                   seed = 123, 
                   init = "0"
                   )

#save models
saveRDS(mod_doy, file = here::here("data-raw/models", paste0("mod_", gsub("_data_subset.csv", "_doy.rds", filename))))
saveRDS(mod_temp, file = here::here("data-raw/models", paste0("mod_", gsub("_data_subset.csv", "_temp.rds", filename))))
}

# List of filenames
filenames <- c("natural_chinook_data_subset.csv", "hatchery_chinook_data_subset.csv", "natural_steelhead_data_subset.csv", "hatchery_steelhead_data_subset.csv")

# Loop over the filenames and call the function for each one
for (filename in filenames) {
  run_models(filename)
}

```

# individual run (if needed)

##subset raw data
Import each subset file and run model for doy and temp. Save to data-raw/models folder

```{r data_import}
#import saved subset file 
df<-read_csv(here("data-raw/HydroSurvDOYTEMP_data_subset", "natural_chinook_data_subset.csv"))
df<-read_csv(here("data-raw/HydroSurvDOYTEMP_data_subset", "hatchery_chinook_data_subset.csv"))
df<-read_csv(here("data-raw/HydroSurvDOYTEMP_data_subset", "natural_steelhead_data_subset.csv"))
df<-read_csv(here("data-raw/HydroSurvDOYTEMP_data_subset", "hatchery_steelhead_data_subset.csv"))

df<-read_csv(file.choose())
```

## aggregate subset data

```{r data_aggregate}
#for binomial left hand trial 
df.agg <- df%>% 
  #calculate # of juveniles--used as total trials in binomial model
  group_by(year, doy, doyz, transport) %>% 
  mutate(doy = as.numeric(doy),
         n = n()) %>% 
  arrange(year, doy, doyz) %>% 
 # mutate(year = as.factor(year)) %>% 
  #set reponse as sum of alive per trials 
  summarize(alive = sum(alive),
            n = mean(n),
            mean.tempz = mean(BON.tempz),
            mean.temp = mean(BON.temp)) %>% 
  ungroup() %>% 
  #include sar based on pit tag returns of observed data-- binned by year, weekly doy (doy.bin), and transport (ROR v Transport)
  arrange(year, doy, doyz, transport) %>% 
  mutate(doy.bin = cut(doy, breaks = seq(90, 160, 10), include.lowest = TRUE)) %>% 
  group_by(year, transport, doy.bin) %>% 
  mutate(doy.median = round(median(doy), 0),
         sar.pit = alive/n) %>% 
         # sar.pit.se = sd(sar.pit)/sqrt(n),
         # sar.pit.lower = mean(sar.pit) - 1.96 * sar.pit.se,
         # sar.pit.upper = mean(sar.pit) + 1.96 * sar.pit.se) %>% 
  ungroup() %>% 
  mutate( date = parse_date_time(x = paste(year, doy), orders = "yj")) %>% 
  mutate(year = as.character(year),
         transport = as.character(transport))

```

## run model
Based on which species and rear type, update the naming for saving files



```{r, set_prior}
bprior <- c(prior(normal(0, 5), class = b), 
            prior(normal(0, 5), class = sd),
            prior(normal(0, 2), class = sd, group = year))

default_prior(alive | trials(n) ~ doyz + I(doyz^2) + transport + doyz:transport + I(doyz^2):transport + (1 + doyz + I(doyz^2) + transport | year), 
                   data = df.agg, 
                   family = binomial(link = "logit"), prior = bprior)
```



```{r mod_DOY_binomial}
mod_natural_chinook_doy <- brm(alive | trials(n) ~ doyz + I(doyz^2) + transport + doyz:transport + I(doyz^2):transport + (1 + doyz + I(doyz^2) + transport | year), 
                   data = df.agg, 
                   family = binomial(link = "logit"), 
                   chains = 3,
                   cores = 3,
                   prior = bprior,
                   file = here("data-raw/models", "mod_natural_chinook_doy")
                   )

#check summary
summary(mod_natural_chinook_doy)
summary(mod_natural_chinook_doy)

# loo(mod_natural_chinook_doy, mod_natural_chinook_doy_prior)

shinystan::launch_shinystan(mod_natural_chinook_doy)
pp_check(mod_natural_chinook_doy)
prior_summary(mod_natural_chinook_doy)
```

```{r mod_TEMP_binomial}

mod_natural_steelhead_temp <- brm(alive | trials(n) ~ mean.tempz + I(mean.tempz^2) + transport + mean.tempz:transport + I(mean.tempz^2):transport + (1 + mean.tempz + I(mean.tempz^2) + transport | year), 
                   data = df.agg, 
                   family = binomial(link = "logit"), 
                   chains = 3,
                   cores = 3,
                   prior = bprior,
                   file = here("data-raw/models", "mod_natural_steelhead_temp")
                   )

#check summary
summary(mod_natural_steelhead_temp)

```


```{r compare, eval= FALSE}
 
r<-loo(mod_natural_chinook_doy, mod_natural_chinook_temp)
kable(r$diffs)

```




