---
title: "Subset original data and append river temperature data"
output: html_notebook
---

This code is was used to produce the subset dataset used for each model which can be found in the data-raw/ folder. 

This data is not necessary to run, but kept for reference.


```{r load_libraries, eval=FALSE}
library(data.table)
library(here)
library(tidyverse)
library(zoo)
```


# Original dataset

Original data sets were shared by JG via baloo link  - included in folder as zip file for reference. Using code below a subset of .csv per species and rear type was created and saved in the data-raw/HydroSurvDOYTEMP_data_subset.zip folder. 

```{r wrangle_orig_data,  eval=FALSE}


# DATA <- read.csv(file = here("Examples/data", "dsynth_2021-03-02.spsu.subset.csv"), stringsAsFactors = FALSE, na.strings = c("", "NA"))
# 
# DATA <- read.csv(file = here("Examples/data", "dsynth_2021-03-02.stl.subset.csv"), stringsAsFactors = FALSE, na.strings = c("", "NA"))

DATA <- read.csv(file = file.choose(), stringsAsFactors = FALSE, na.strings = c("", "NA")) #select either spsu  ch or stl subset file

df<- DATA %>%
  # filter to include wild all run types, with a transport code released from LGR
  filter(
    t_rear_type == "W",
    !is.na(pass_type),
    str_detect(rel_site, "LGR")
  ) %>%
  # extract year/doy and filter to include 80 to 160 doy from 1993 to 2018
  mutate(
    year = lubridate::year(rel_time),
    doy = lubridate::yday(rel_time),
  ) %>%
  
  filter(
    between(year, 1993, 2018),
    between(doy, 80, 160)
  ) %>%
  # redesignate tranportation codes--double check on -S and -T at other locations 
  mutate(pass_type_T_R = case_when(
    str_detect(pass_type, "-S") & str_length(pass_type) == 5 ~ "remove",
    str_detect(pass_type, "-S") & str_length(pass_type) > 5 ~ "ROR",
    str_detect(pass_type, "-TB") ~ "ROR",
    str_detect(pass_type, "-TD") ~ "ROR",
    str_detect(pass_type, "ROR") ~ "ROR",
    str_detect(pass_type, "LWG-T") ~ "T",
    .default = "remove"
  )) %>%
  filter(pass_type_T_R != "remove") %>% 
  # # adding filters to match JG shared code for transport --- check this still stands
  # filter(if_else(pass_type_T_R == "ROR", 
  #                str_length(pass_type) %in% c(3, 7),
  #                str_detect(pass_type, "LWG-T")
  #                )
  #        ) %>%
  # add binary codes
  mutate(
    transport = if_else(pass_type_T_R == "T", 1, 0),
    alive = if_else(!is.na(adu_lgr_first), 1, 0),
    doyz = scale(doy, center = TRUE, scale = TRUE)
  ) 

#save subset output to data-raw/ folder --rerun and update naming for each subsetted dataset
#write.csv(df, row.names = F, here("data-raw", "wild_spsu_chinook_data_subset.csv"))

```

# River data

This code is used to import BON river temperature data and calculate 7-day rolling averages for the years 1993-2018. The data is then merged into a larger dataset for further analysis. 


```{r setup, eval=FALSE}
BON.temp.raw<-read_csv(here::here("Examples/data", "rivermg_1704223998_133.csv")) #update to pull from CBR link

BON.temp.data <- BON.temp.raw %>%
  # mutate(across(everything(), ~ifelse(.=="", NA, as.character(.)))) %>%
    filter(!row_number() %in% c(367:370)) %>% 
  pivot_longer(
    cols = c(2:27),
    names_to = "ID",
    values_to = "value"
  ) %>%
  separate(ID, c("CY", "BON", "ID")) %>%
  pivot_wider(
    names_from = "ID",
    values_from = "value"
  ) %>%
  mutate_at(c("tempc"), as.numeric) %>%
  separate(1, c("mm", "dd")) %>%
  mutate(
    CY = as.numeric(CY),
    mm = as.numeric(mm),
    dd = as.numeric(dd)
  ) %>%
  arrange(CY) %>% # sort data to use rollmean
  group_by(CY) %>%
  filter(complete.cases(tempc)) %>% #remove any days without temp (following JG's code) -update for age of return?
  mutate(
    BON.temp = rollmean(tempc, k = 7, fill = NA, align = "right"), #same as JG code, returns NA for 7 day average with NA's present
    BON.temp.apply = rollapply(tempc, width = 7, FUN = function(x) mean(x, na.rm = TRUE), by = 1, by.column = TRUE, partial = TRUE, fill = NA, align = "right") #adjusts for NA's within d7 averages
  ) %>%
  mutate(YMD = lubridate::ymd(paste(CY, mm, dd, sep = "-"))) %>%
  mutate(DOY = lubridate::yday(YMD)) %>%
 # select("CY","mm", "DOY", "BON.temp") %>%
  mutate(
    CY = as.character(CY),
    DOY = as.character(DOY)
  )




df %>% 
  left_join(BON.temp.data, by = c("year" = "CY", "doy" = "DOY")) 

  #write_csv(here::here("data-raw", "X_X_data_subset.csv"))


```

```{r bind_temp_data, eval=FALSE}

#merge with subset dataset and save to data-raw/ folder
df<-file.choose() #choose the subsetted dataset to merge with river temperature data


df<- df%>% 
  mutate(
    year = as.character(year),
    doy = as.character(doy)
  ) %>% 
  left_join(y= select(BON.temp.data, BON.temp, BON.temp.apply, CY, DOY), by = c("year" = "CY", "doy" = "DOY")) %>% 
  mutate ( BON.tempz = scale(BON.temp, center = TRUE, scale = TRUE)
  ) %>% 
  select(!c( BON.temp.apply))

# save subset output to data-raw/ folder and update naming
#write.csv(df, row.names = F, here("data-raw/HydroSurvDOYTEMP_data_subset", "wild_steelhead_data_subset.csv"))
```

