---
title: "Posterior predict options"
date: "2023-12-13"
output: 
  html_document:
    code_folding: show
---
Below outlines code comparing posterior predictions using tidybayes:

**Data wrangling**

Data & models used:
```{r data_model_packages_import, class.source="fold-hide", warning=FALSE, message=FALSE, error=FALSE}
library(tidybayes)
library(brms)
library(tidyverse)
library(dplyr)
library(here)
library(knitr)
library(kableExtra)

data<-read.csv(here::here("Examples/data", "ChSSWRT_data_subset.csv"))

mod.ChSSWRT <- brms::brm(alive | trials(n) ~ doyz + I(doyz^2) + transport + doyz:transport + I(doyz^2):transport + (1 + doyz + I(doyz^2) + transport | year),
                   data = df.agg, 
                   family = binomial(link = "logit"), 
                   chains = 3,
                   cores = 3,
                   file = here::here("Examples/results", "mod.wild_Chinook_DOY.rds"))
```

Aggregating data in form (alive | trial) for binomial model
```{r data_aggregate, class.source='fold-hide', warning=FALSE, message=FALSE}
#for binomial left hand trial 
df.agg <- data %>% 
  #calculate # of juveniles--used as total trials in binomial model
  group_by(year, doy, doyz, transport) %>% 
  mutate(n = n()) %>% 
  arrange(year, doy, doyz) %>% 
  #set reponse as sum of alive per trials 
  summarize(alive = sum(alive),
            n = mean(n)) %>% 
  ungroup() %>% 
  #include sar based on pit tag returns of observed data-- binned by year, weekly doy (doy.bin), and transport (ROR v Transport)
  arrange(year, doy, doyz, transport) %>% 
  mutate(doy.bin = cut(doy, breaks = seq(90, 160, 10), include.lowest = TRUE)) %>% 
  group_by(year, transport, doy.bin) %>% 
  mutate(doy.median = round(median(doy), 0),
         sar.pit = alive/n ) %>% 
  ungroup() %>% 
  mutate( date = parse_date_time(x = paste(year, doy), orders = "yj"))
```

**Predictions:**

Question: For new data used to predict, should I use a mean number of trials per species and rear_type, or a range of values?
```{r create_newdata}
##create new dataframe with all transport and years, currently set n trials to mean per species & rear_type 

scale.DOY <- scale(data$doy) 
scale.x <- attr(scale.DOY, "scaled:center")
scale.sd <- attr(scale.DOY, "scaled:scale")

#reassigns scale used for model to 90:160 for new dataset used in predictions. 
newdoyz <- ((90:160) - scale.x) / scale.sd 


newdata <- expand_grid(
  doyz = newdoyz,
  transport = c(0, 1),
  year = c(1993:2018),
  n = round(mean(df.agg$n), 0)  #could use range, but currently set to mean
  ) 

```

Import previously used code for glmm:
```{r JG_Code, class.source="fold-hide"}
load(here("Examples/results", "glmm7.sal"))

newdata.jg<-newdata %>% 
  rename(DOYz = doyz,
         Transport = transport,
         Year = year)

glmm_predict <- predict(glmm7.sal, newdata.jg, re.form = ~ (1 + DOYz + I(DOYz^2) + Transport | Year), allow.new.levels = TRUE)


##transform and create df
df.glmm.pred <- data.frame(newdata.jg, glmm_predict, plogis(glmm_predict))

summary_glmm_pred <- df.glmm.pred |> 
  ungroup() |> 
  summarize(across(plogis.glmm_predict., lst(mean, sd, median), .names = "{.fn}"))

```


Compare predictions using tidybayes `linpred_draws()`, `epred_draws()`, and `predicted_draws()`. `epred_draws()`, and `predicted_draws()` produce draws or predicted # of alive per trial. Each of those is then further calculated to get probability whereas `linpred_draws()` is in probability scale and just needs to be transformed. 
```{r compare_posteriorpredictions, class.source='fold-hide'}
#differences in linpred, epred, pred using tidybayes https://www.andrewheiss.com/blog/2022/09/26/guide-visualizing-types-posteriors/

#predict
df.explinpred<-newdata %>% 
  add_linpred_draws(mod.ChSSWRT, re_formula = NULL, allow_new_levels = TRUE, ndraws = 10,  transform = TRUE, value = ".explinpred") 

df.epred<-newdata %>% 
  add_epred_draws(mod.ChSSWRT, re_formula = NULL, allow_new_levels = TRUE, ndraws = 10) %>% 
    mutate(.probability = .epred/n)
 
df.pred<-newdata %>%  
   add_predicted_draws(mod.ChSSWRT, re_formula = NULL, allow_new_levels = TRUE, ndraws = 10) %>% 
  mutate(.probability = .prediction/n)

#summarize

summary_exp_linpred <- df.explinpred |> 
  ungroup() |> 
  summarize(across(.explinpred, lst(mean, sd, median), .names = "{.fn}"))

summary_normal_epred <- df.epred |> 
  ungroup() |> 
  summarize(across(.probability, lst(mean, sd, median), .names = "{.fn}"))

summary_normal_predicted <- df.pred |> 
  ungroup() |> 
  summarize(across(.probability, lst(mean, sd, median), .names = "{.fn}"))
```


```{r table_posteriorpredictions, echo=FALSE}
tribble(
  ~Function,
  "<code>predict(glmm, re_formula = NA)</code>",
  "<code>posterior_linpred(re_formula = NULL, transform = TRUE)</code>",
   "<code>posterior_epred(re_formula = NULL)</code>", 
    "<code>posterior_predict(re_formula = NULL)</code>", 
) |> 
  bind_cols(bind_rows(summary_glmm_pred, summary_exp_linpred, summary_normal_epred,summary_normal_predicted)) |> 
  kbl(escape = FALSE) |> 
  kable_styling()

```




Plots of each comparison:

```{r compare_plots_allyears, echo=FALSE, out.width = "50%"}


df.explinpred %>% 
tidybayes::median_qi() %>% 
ggplot(  aes(x=doyz, y=.explinpred, group = as.factor(transport),  color= as.factor(year)))+
        labs(title = "posterior_linpred( re_formula = NULL, transform = TRUE)" ) +
  geom_point()


df.epred %>% 
tidybayes::median_qi(.probability) %>% 
ggplot(  aes(x=doyz, y=.probability, group = as.factor(transport),  color= as.factor(year)))+
        labs(title = "posterior_epred( re_formula = NULL)" ) +
  geom_point()


df.pred %>% 
tidybayes::median_qi(.probability) %>% 
ggplot(  aes(x=doyz, y=.probability, group = as.factor(transport),  color= as.factor(year)))+
      labs(title = "posterior_pred( re_formula = NULL)" ) +
  geom_point()

df.glmm.pred %>% 
ggplot(  aes(x=DOYz, y=plogis.glmm_predict., group = as.factor(Transport),  color= as.factor(Year)))+
    labs(title = "predict(glmm, re_formula = NULL)" ) +
  geom_point()


```

